---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data
```{r}
unzip("~/RepData_PeerAssessment1/activity.zip") 
activity_data<-read.csv("~/RepData_PeerAssessment1/activity.csv")

```


## What is mean total number of steps taken per day?
1.Calculate the total number of steps taken per day
```{r}
per_day_steps<-tapply(activity_data$steps,activity_data$date,sum,na.rm=T)
per_day_steps
```
2.Make a histogram of the total number of steps taken each day
```{r}
hist(per_day_steps)
```

3.Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean_per_day_steps<-mean(per_day_steps)
median_per_day_steps<-median(per_day_steps)
mean_per_day_steps
median_per_day_steps
```


## What is the average daily activity pattern?
1.Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
avg_per_interval_steps<-aggregate(activity_data$steps,list(activity_data$interval),mean,na.rm=T)
names(avg_per_interval_steps)<-c("interval","avg_steps")
plot(avg_per_interval_steps$interval,avg_per_interval_steps$avg_steps,type="l",main="Average Number of Steps Per Interval Across All Days")
```

2.Report the 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps.
```{r}
interval_of_max_steps<-avg_per_interval_steps[which(avg_per_interval_steps$avg_steps==max(avg_per_interval_steps$avg_steps)),]$interval
interval_of_max_steps
```


## Imputing missing values
1.Calculate and report the total number of missing values in the dataset 
```{r}
is_NA_activity<-activity_data[is.na(activity_data),]
nrow(is_NA_activity)
```
2.Devise a strategy for filling in all of the missing values in the dataset using the mean for the 5-minute interval.
```{r}
library(dplyr)
temp<-left_join(is_NA_activity,avg_per_interval_steps,by="interval")
```
3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activity_data2<-activity_data
activity_data2[is.na(activity_data2),]$steps<-temp$avg_steps
```
4.Make a histogram of the total number of steps taken each day. Calculate and report the mean and median total number of steps taken per day. 
```{r}
per_day_steps2<-tapply(activity_data2$steps,activity_data2$date,sum,na.rm=T)
hist(per_day_steps2)
mean_per_day_steps2<-mean(per_day_steps2)
median_per_day_steps2<-median(per_day_steps2)
mean_per_day_steps2
median_per_day_steps2
```
From the calculation above, we see that both the mean and median total number of steps taken per day have increased to 10766.19.



## Are there differences in activity patterns between weekdays and weekends?
1.Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
activity_data2$date<-as.Date(activity_data2$date)
activity_data2$weekday<-as.factor(weekdays(activity_data2$date))
levels(activity_data2$weekday)<-c("Weekday","Weekday","Weekend","Weekend","Weekday","Weekday","Weekday")
```
2.Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
avg_per_interval_steps2<-aggregate(activity_data2$steps,list(activity_data2$interval,activity_data2$weekday),mean,na.rm=T)
names(avg_per_interval_steps2)<-c("interval","weekday","avg_steps")
library(ggplot2)
ggplot(avg_per_interval_steps2,aes(x=interval,y=avg_steps))+facet_grid(weekday~.)+geom_line()
```

From the plot above we can see that the activity pattern in weekdays are different from the pattern in weekends. The former has only one peak and higher average number of steps per interval while the latter has multiple peaks and 